### Read activity.csv dataset into activity

```{r, echo=TRUE}
activity <- read.csv("C:/Users/dheer/Desktop/Coursera/5. Reproducible Research/project/activity.csv", stringsAsFactors = FALSE)
str(activity)
```


```{r, echo=TRUE}
# date is converted to date dattype,
activity$date <- as.Date(activity$date)
#Total number of steps per day are calculated and histogram is plotted
total_steps <- aggregate(steps~date, activity, sum, na.rm = TRUE)
hist(total_steps$steps, breaks = 20, col = "red", xlab = "Total_Steps", 
                         main = "Histogram of total number of steps each day")
```

Mean number of steps taken each day
```{r, echo=TRUE}
total_mean <- mean(total_steps$steps)
total_mean
```

Mean number of steps taken each day
```{r,echo = TRUE}
total_median <- median(total_steps$steps)
total_median
```

Time series plot of the average number of steps taken
```{r, echo=TRUE}
avg_steps <- aggregate(steps~interval, activity, mean, na.rm = TRUE)

plot(avg_steps$interval,avg_steps$steps,type= "l",xlab = "Interval",ylab = "Average Steps", 
     main = "Time Series plot of the average number of steps taken")
```

The 5-minute interval that, on average, contains the maximum number of steps
```{r,echo = TRUE}
avg_steps[which.max(avg_steps$steps),]
```

Code to describe and show a strategy for imputing missing data
```{r,echo = TRUE}
Imp_activity <- activity
count <- sum(is.na(Imp_activity$steps))
print("Number of NA values:")
count
Imp_activity$steps[is.na(Imp_activity$steps)] <- mean(Imp_activity$steps,na.rm = TRUE)
count <- sum(is.na(Imp_activity$steps))
print("Number of NA values after imputing:")
count
```

```{r,echo = TRUE}
total_steps_imput <- aggregate(steps~date, Imp_activity, sum, na.rm = TRUE)
hist(total_steps_imput$steps, breaks = 20, col = "red", xlab = "Total_steps",
     main = "Histogram of the total number of steps taken each day after imputing")
```

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r,echo = TRUE}
Imp_activity$week <- ifelse (weekdays(Imp_activity$date,abbreviate = FALSE) %in% c("Saturday","Sunday"), 
                         "Weekend", "Weekday")
avg_steps_new <- aggregate(steps~interval+week, Imp_activity, mean)
library(lattice)
xyplot(avg_steps_new$steps~avg_steps_new$interval|avg_steps_new$week,type= "l",  layout = c(1,2), xlab = "Interval",
       ylab = "Avg_steps",
       main = "Average steps taken per 5-minute interval across weekdays and weekends")
```

```{r,echo = TRUE}
if(!file.exists("./Figure")){dir.create("./Figure")}
png("./plot1.png")
hist(total_steps$steps, breaks = 20, col = "red", xlab = "Total_Steps", 
                         main = "Histogram of total number of steps each day")
dev.off()

png("./plot2.png")
plot(avg_steps$interval,avg_steps$steps,type= "l")
dev.off()

png("./plot3.png")
hist(total_steps_imput$steps, breaks = 20, col = "red", xlab = "Total_steps",
     main = "Histogram of the total number of steps taken each day after imputing")
dev.off()

png("./plot4.png")
xyplot(avg_steps_new$steps~avg_steps_new$interval|avg_steps_new$week,type= "l",  layout = c(1,2), xlab = "Interval",
       ylab = "Avg_steps",
       main = "Average steps taken per 5-minute interval across weekdays and weekends")
dev.off()